---
title: "RNAi_process"
author: "Patricka Williams-Simon & Libby King"
date: "4/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DESeq2)
library(dplyr)
library(stringr)
library(sva)
library(ggplot2)
library(cowplot)
library(colorspace)
library(ggrepel)
library(tidyverse)
library(multcomp)
theme_set(theme_cowplot())

source(file="../../../../../kingeg/Functions/ggplot_theme.R")
```

## Read in processed data

```{r}

ThermDat <- readRDS(file="../ProcessedData/Incap_processed.Rds")
str(ThermDat)
ThermDat[1:5,]

#add a col set id 
ThermDat$set_id<- paste(ThermDat$genotype, ThermDat$cross, sep="_")

RNAi_backg <- read_excel("../ProcessedData/HeatPlate_autotrack/RNAi_Bckg.xlsx")
str(RNAi_backg)

colnames(RNAi_backg) <-c("genotype", "backg_id", "Gname")

ThermDat <- left_join(ThermDat, RNAi_backg, by="genotype")
```

## Data Checking

Use `unique()` to check groups, replicates, genotypes, etc. for issues. You will need a column that specifies which flies you want to compare so work toward getting that set up here and checking for any problems. Use `tally()` to count up how many you have in your groups to make sure that makes sense too. Etc.



```{r}

#removed individuals with incap time of 5 secs or less

ThermDat<- ThermDat %>% filter(incapacitation > 5)

p1<- ggplot(ThermDat, aes(x = incapacitation)) +
  geom_histogram()


#check number of individuals per genotype 
RNAicounts <- ThermDat%>% group_by(set_id) %>% tally()

hist(RNAicounts$n)

#look at genotypes that have less that 30 and more than 50 individuals 
RNAicounts[which(RNAicounts$n < 30),]
RNAicounts[which(RNAicounts$n > 40),]


#cleaning up data. See notes below
ThermDat[which(ThermDat$set_id=="60071_25752"), 'cross'] <-25750
ThermDat[which(ThermDat$set_id=="33545_3954"), 'genotype'] <-33646

ThermDat <- ThermDat %>% filter(!(set_id %in% c("32913_25750", "32913_3954", "32913_25750", "50656_51941")))

str(ThermDat)
ThermDat$set_id<- paste(ThermDat$genotype, ThermDat$cross, sep="_")


unique(ThermDat$group)
unique(Gene$set_id)
unique(ThermDat$genotype)

#sort data from min to max
RNAiData_sort <- ThermDat[order(ThermDat$incapacitation),]

RNAiData_sort[1:5,]
```

Dropped all lines with genotype 32913 because it didn't work well and we had an incomplete set.
Dropped 50656 because there was only 2 individuals measured for TT. 
Merged 60071-25750 and 60071-25752 because there was a typo in file naming and 60071-25752 is not a genotype. The file name column did not change. 
Merged 33545_3954 and 33646_3954  because there was a typo in file naming and 33545_3954 is not a genotype.


#calculate mean and SEM
```{r}
RNAi_Data <- ThermDat %>% 
  group_by(set_id) %>% 
  summarise(Incap_Mean = mean(incapacitation),
            Incap_SEM = sd(incapacitation)/sqrt(length(incapacitation)),
            N = length(incapacitation))

which.min(RNAi_Data$N)
mean(RNAi_Data$N)

#Visulalize distribution with histogram for both mean and raw 
RNAi_histo <- ggplot(RNAi_Data, aes(x = Incap_Mean)) +
  geom_histogram()

RNAi_Datahisto <- ggplot(ThermDat, aes(x = incapacitation)) +
  geom_histogram()

```

#running Model

For each gene compare to Gal4 crosses and background (P2 or P40)


```{r}

#get vector of focal genes
ThermDat[1:5,]

Glist<- unique(ThermDat$genotype)
str(Glist)

#remove background lines  from Glist 
Glist<- Glist[which(!(Glist %in% c("36303", "36304")))]


nn <- numeric(length=length(Glist))
counter <- 1

for(foc.gene in Glist)
  {
  
#foc.gene <- 31167
Gene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(Gene$backg_id[1]=="attP2",36303,36304))

Gene <- rbind(Gene, Bk)

RNAi_mod <- lm(incapacitation ~ set_id, data= Gene)
 summary(RNAi_mod)

 str(RNAi_mod)
 
 aa<-aov(RNAi_mod)
nn[counter]<- aa[1,5]
 Gene %>% 
   group_by(set_id) %>%
   dplyr::summarise(mean(incapacitation))
 counter<-counter+1
 
}

ThermDat[1:5,]




#adusted p-value 
rr <- p.adjust(nn, method="BH")


#finding significance for three genes within group 
#RhoGAP93B
foc.gene <- 35027
Gene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(Gene$backg_id[1]=="attP2",36303,36304))

Gene <- rbind(Gene, Bk)

RNAi_mod <- lm(incapacitation ~ set_id, data= Gene)
 summary(RNAi_mod)

 str(RNAi_mod)
 
 aa<-aov(RNAi_mod)

#post-hoc test 
Rho<-TukeyHSD(aa)
plot(Rho)


#dmrt93B
foc.gene <- 50656
Gene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(Gene$backg_id[1]=="attP2",36303,36304))

Gene <- rbind(Gene, Bk)

RNAi_mod <- lm(incapacitation ~ set_id, data= Gene)
 summary(RNAi_mod)

 str(RNAi_mod)
 
 aa<-aov(RNAi_mod)

#post-hoc test 
dmrt<-TukeyHSD(aa)
plot(dmrt)


#atpalpha
foc.gene <- 51411
Gene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(Gene$backg_id[1]=="attP2",36303,36304))

Gene <- rbind(Gene, Bk)

RNAi_mod <- lm(incapacitation ~ set_id, data= Gene)
 summary(RNAi_mod)

 str(RNAi_mod)
 
 aa<-aov(RNAi_mod)

#post-hoc test 
atp<-TukeyHSD(aa)
plot(atp)


#merge padj to Glist 
Glist <- cbind(Glist, rr)


Glist <- as_data_frame(Glist,Glist = "unique")
str(Glist)
```


Visualizing data 
```{r}
Gname<- unique(ThermDat$Gname)

Glist <-cbind(Glist, Gname)



foc.gene <- 51411
FGene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(FGene$backg_id[1]=="attP2",36303,36304))

FGene <- rbind(FGene, Bk)


Atpha_RNAi_plot <- ggplot(FGene, aes(set_id, incapacitation, color=set_id )) +
  geom_point(position = position_jitter(width = 0.4), alpha = 1/4) +
  stat_summary(fun = mean, geom = "point", pch = 12, size = 5) +
  stat_summary(fun.data = mean_se, 
               geom = "errorbar", 
               color = 'black', width=.3, size=0.5, alpha=1/2) +
  ggtitle("Na pump α subunit (atpα)") +
  guides(col=guide_legend("Group Mean ")) +
  ylab("Incapacitation Time (sec)") +
  xlab("RNAi Line") 
  #my_theme


foc.gene <- 50656
FGene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(FGene$backg_id[1]=="attP2",36303,36304))

FGene <- rbind(FGene, Bk)


dmr_RNAi_plot <- ggplot(FGene, aes(set_id, incapacitation, color=set_id )) +
  geom_point(position = position_jitter(width = 0.4), alpha = 1/4) +
  stat_summary(fun = mean, geom = "point", pch = 12, size = 5) +
  stat_summary(fun.data = mean_se, 
               geom = "errorbar", 
               color = 'black', width=.3, size=0.5, alpha=1/2) +
  ggtitle("doublesex-Mab related 93B (dmr938)") +
  guides(col=guide_legend("Group Mean ")) +
  ylab("Incapacitation Time (sec)") +
  xlab("RNAi Line") 
  #my_theme


foc.gene <- 35027
FGene <- ThermDat %>% filter(genotype ==foc.gene)

Bk <- ThermDat %>% filter(genotype == ifelse(FGene$backg_id[1]=="attP2",36303,36304))

FGene <- rbind(FGene, Bk)


Rho_RNAi_plot <- ggplot(FGene, aes(set_id, incapacitation, color=set_id )) +
  geom_point(position = position_jitter(width = 0.4), alpha = 1/4) +
  stat_summary(fun = mean, geom = "point", pch = 12, size = 5) +
  stat_summary(fun.data = mean_se, 
               geom = "errorbar", 
               color = 'black', width=.3, size=0.5, alpha=1/2) +
  ggtitle("Rho GTPase activating protein at 93B (RhoGAP93B)") +
  guides(col=guide_legend("Group Mean ")) +
  ylab("Incapacitation Time (sec)") +
  xlab("RNAi Line") 
  #my_theme


```