---
title: "RNAi_Sig_Genes"
author: "Patricka Williams-Simon"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script is to analyse a subset of RNAi genes that was re-phenotyped for thermal tolerance.
## We refer to this dataset as RNAi2

## load file 
```{r cars}
ThermDat2 <- read.csv(file="../ProcessedData/RNAi2_tracked_data.csv")
str(ThermDat2)
ThermDat2[1:5,]
ThermDat2$genotype  <- as.character(ThermDat2$genotype)

#add a col set id 
ThermDat2$set_id<- paste(ThermDat2$genotype, ThermDat2$cross, sep="_")


#load dataset with background/ crosses information
RNAi_backg <- read_excel("../ProcessedData/HeatPlate_autotrack/RNAi_Bckg.xlsx")
str(RNAi_backg)
RNAi_backg[1:5,]

#assign column names
colnames(RNAi_backg) <-c("genotype", "backg_id", "Gname")

#combine data sets
ThermDat2 <- left_join(ThermDat2, RNAi_backg, by="genotype")

```

## Data Checking

Use `unique()` to check groups, replicates, genotypes, etc. for issues. You will need a column that specifies which flies you want to compare so work toward getting that set up here and checking for any problems. Use `tally()` to count up how many you have in your groups to make sure that makes sense too. Etc.


```{r}


#removed individuals with incap times of 5 secs or less
ThermDat2<-ThermDat2 %>% filter(incapacitation > 5)
str(ThermDat2)

#Visualize after filtering 
p1<- ggplot(ThermDat2, aes(x = incapacitation)) +
  geom_histogram()


#look at the individuals that have scores higher than 300 secs
ThermDat2[which(ThermDat2$incapacitation > 300),]


#check number of individuals per genotype 
RNAicounts2 <- ThermDat2%>% group_by(set_id) %>% tally()

#visualize 
hist(RNAicounts2$n)


#identify crosses with more/ less individuals than expected
RNAicounts2[which(RNAicounts2$n < 25),]


#check for duplicates in data set
dupsRNAi2 <- which(duplicated(ThermDat2, MARGIN=1, incomparables = FALSE ))
RNAi2_ch[which(duplicated(RNAi2_ch$file)),]


save(dupsRNAi2, file="../ProcessedData/DupsRNAi2.rda")


#read in data sheet with expected pheno total
RNAi2_ch <- read_excel("../ProcessedData/HeatPlate_autotrack/RNAi_Val2_Chmb_positions.xlsx")
RNAi2_ch$file <- paste0(RNAi2_ch$FileName,"_", RNAi2_ch$Chamber)

#confirm where files were duplicated files from top to bottom
which(duplicated(RNAi2_ch$file))

# confirm where files were duplicated files from bottom to top
which(duplicated(RNAi2_ch$file, fromLast=TRUE))
RNAi2_ch[which(duplicated(RNAi2_ch$file)),]

#remove duplicated rows 
ThermDat2<- ThermDat2[-c(171), ]

#confirm duplicates were removed, if removed the output should read integer(0) 
dupsRNAi2 <- which(duplicated(ThermDat2, MARGIN=1, incomparables = FALSE ))


#sort data from min to max
RNAiData_sort2 <- ThermDat2[order(ThermDat2$incapacitation),]


#look at distrubution of data
hist(ThermDat2$incapacitation)
str(ThermDat2)

#identify outliers
RNAiData_sort2[1:10,]

```

#running Model

For each gene compare to Gal4 crosses and background (P2 or P40)

```{r}

#get vector of focal genes
ThermDat2[1:5,]


Glist2<- unique(ThermDat2$genotype)
str(Glist2)

#remove background lines  from Glist 
Glist2<- Glist2[which(!(Glist2 %in% c("36303", "36304")))]

nn2 <- numeric(length=length(Glist2))

counter <- 1


#performing lm and ANOVA in a for loop

for(foc.gene in Glist2)
  {
  
Gene2 <- ThermDat2 %>% filter(genotype ==foc.gene)
Gene2[1:5,]

Bk <- ThermDat2 %>% filter(genotype == ifelse(Gene2$backg_id.x[1]=="attP2",36303,36304))

Gene2 <- rbind(Gene2, Bk)

RNAi_mod2 <- lm(incapacitation ~ set_id, data= Gene2)
 summary(RNAi_mod2)

 str(RNAi_mod2)
 
 aa2<-aov(RNAi_mod2)
 pp<- summary(aa2)
 str(pp)
 pp<-pp[[1]] 
 pp[1,5]
 
nn2[counter]<- pp[1,5]
 
Gene2 %>% 
   group_by(set_id) %>%
   dplyr::summarise(mean(incapacitation))
 counter<-counter+1
 
}

#view which genes are significant
nn2

#post-hoc test 
aa2_sig<-TukeyHSD(aa2)
plot(aa2_sig)

#adusted p-value 
rr <- p.adjust(nn2, method="BH")


#merge files 
Glist2 <- cbind(Glist2, rr)


```

###Visualizing data 
```{r}

foc.gene <- 51411
FGene2 <- ThermDat2 %>% filter(genotype ==foc.gene)

Bk <- ThermDat2 %>% filter(genotype == ifelse(FGene2$backg_id.x[1]=="attP2",36303,36304))

FGene2 <- rbind(FGene2, Bk)


Atpha_RNAi_plot2 <- ggplot(FGene2, aes(set_id, incapacitation, color=set_id )) +
  geom_point(position = position_jitter(width = 0.4), alpha = 1/4) +
  stat_summary(fun = mean, geom = "point", pch = 12, size = 5) +
  stat_summary(fun.data = mean_se, 
               geom = "errorbar", 
               color = 'black', width=.3, size=0.5, alpha=1/2) +
  ggtitle("Na pump α subunit (atpα) - 2") +
  guides(col=guide_legend("Group Mean ")) +
  ylab("Incapacitation Time (sec)") +
  xlab("RNAi Line") 
  #my_theme



```