---
title: "Processing_Phenotype_Data"
author: "Patricka Williams-Simon"
date: "7/13/2018"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(ggplot2)
library(DSPRqtl)
library(DSPRqtlDataA)
library(tidyverse)
```

## Read in LearnMem processed raw data and check 

```{r}
#rawd <- read.table("/MyGitHub/LearnMemTol/Data/HeatProc.txt",sep="\t", #header=TRUE, stringsAsFactors = FALSE)

rawdLM <- read.table("../Data/HeatProc_Learn.txt",sep="\t", header=TRUE, stringsAsFactors = FALSE)


str(rawdLM)
rawdLM[1:5]

```

## Read in ThermTol processed raw data and check
```{r}
rawdTT <- read.table("../Data/ThermalTol_processed.txt",sep="\t", header=TRUE, stringsAsFactors = FALSE)


str(rawdTT)
rawdTT[1:5]


```

## Find means and SEMs for all three traits

```{r}
L_MDATA <- rawdLM %>% 
  group_by(patRIL) %>% 
  summarise(Learning_Mean = mean(Learning),
            Learning_SEM = sd(Learning)/sqrt(length(Learning)),
            Memory_Mean = mean(Memory),
            Memory_SEM = sd(Memory)/sqrt(length(Memory)),
            N = length(Memory))




T_TDATA <- rawdTT %>% 
  group_by(patRIL) %>% 
  summarise(Incapacitation_Mean = mean(incapacitation),
            Incapacitation_SEM = sd(incapacitation)/sqrt(length(incapacitation)),
            N = length(incapacitation))
```

## Check to see if SEMs are SEMs (optional)

```{r}
L_MDATA[1:10,c('patRIL','Learning_SEM')]

#sd(rawd[rawd$patRIL==11002,'Learning'])/sqrt(length(rawd[rawd$patRIL==11002,'Learning']))


```

#Find Percentile for each RIL

```{r}

L_MDATA$Percentile_Learning <- ecdf(L_MDATA$Learning_Mean)(L_MDATA$Learning_Mean)

L_MDATA$Percentile_Memory <- ecdf(L_MDATA$Memory_Mean)(L_MDATA$Memory_Mean)

T_TDATA$Percentile_ThermTol<- ecdf(T_TDATA$Incapacitation_Mean)(T_TDATA$Incapacitation_Mean)

```


## Save data (note to self: make sure working directory is set as ~/MyGitHUb/LearnMemTol/Data/)

```{r}
save(L_MDATA, file="../Data/L_MDATA.rda")

save(T_TDATA, file="../Data/T_TDATA.rda")

```

## Visualize data

```{r}

load(file="../Data/L_MDATA.rda")
load(file="../Data/T_TDATA.rda")


#Make histogram LearnMem
Learnhisto <- ggplot(L_MDATA, aes(x = Learning_Mean)) +
  geom_histogram()

Memhisto <- ggplot(L_MDATA, aes(x = Memory_Mean)) +
  geom_histogram()

ThermTolhisto <- ggplot(T_TDATA, aes(x = Incapacitation_Mean)) +
  geom_histogram()


#View graph
Learnhisto

Memhisto

ThermTolhisto

```

## Check dataset

```{r}
#Identify minimum mean (find really low learning point identified in histograms)
L_MDATA[which.min(L_MDATA$Learning_Mean),]
L_MDATA[which.min(L_MDATA$Memory_Mean),]
T_TDATA[which.min(T_TDATA$Incapacitation_Mean),]


#Look at indivial Learning PIs for RILs with lowest means
rawdLM[rawdLM$patRIL==12114,'Learning']
rawdLM[rawdLM$patRIL==11189,'Memory']

#Look at indivial Incapacitation times for RIL with lowest mean
rawdTT[rawdTT$patRIL==11459,'incapacitation']


#Identify maximum means.
L_MDATA[which.max(L_MDATA$Learning_Mean),]
L_MDATA[which.max(L_MDATA$Memory_Mean),]
T_TDATA[which.max(T_TDATA$Incapacitation_Mean),]


#Sort and save from decreasing to increasing means 
LearnDATA_sort <- L_MDATA[order(L_MDATA$Learning_Mean),]
MemDATA_sort <- L_MDATA[order(L_MDATA$Memory_Mean),]
ThermTolDATA_sort <- T_TDATA[order(T_TDATA$Incapacitation_Mean),]

#save sorted dataset
save(LearnDATA_sort, file="../Data/LearnDATA_sort.rda")
save(MemDATA_sort, file="../Data/MemDATA_sort.rda")
save(ThermTolDATA_sort, file="../Data/ThermTolDATA_sort.rda")

#load files, if necessary
load(file ="Data/LearnDATA_sort.rda")
load(file ="Data/MemDATA_sort.rda")
load(file ="Data/ThermTolDATA_sort.rda")

#Identify RILs in the low cohort
LearnDATA_sort[1:20,]
MemDATA_sort[1:20,]
ThermTolDATA_sort[1:20,]


#Identify RILs in the mid cohort
LearnDATA_sort[349:374,]
MemDATA_sort[349:374,]
ThermTolDATA_sort[349:374,]

#Identify RILs in the high cohort
LearnDATA_sort[7:741,]
MemDATA_sort[721:730,]
ThermTolDATA_sort[721:730,]

```


#Identify where RILs used for RNAseq are in the final Dataset

```{r}
#Learning Low
LearnRNAseqLow <- L_MDATA %>% filter(patRIL %in% c("12044", "12285", "12138", "11189", "12022", "12084", "12324"))

save(LearnRNAseqLow, file="Data/LearnRNAseqLow.rda")

#Learning High
LearnRNAseqHigh <- L_MDATA %>% filter(patRIL %in% c("11065", "12359", "11146", "11144", "11006", "11122", "11023"))

save(LearnRNAseqHigh, file="Data/LearnRNAseqHigh.rda")

#Memory Low
MemRNAseqLow <- L_MDATA %>% filter(patRIL %in% c("12030", "12383", "12285", "11189", "11038", "11133", "11186"))

save(MemRNAseqLow, file="Data/MemRNAseqLow.rda")

#Memory High
MemRNAseqHigh <- L_MDATA %>% filter(patRIL %in% c("12018", "12039", "12074", "11184", "11122", "12071", "11023"))

save(MemRNAseqHigh, file="Data/MemRNAseqHigh.rda")


#Thermal Tolerance Low
ThermRNAseqLow <- T_TDATA %>% filter(patRIL %in% c("12047", "12290", "12042", "11106", "12060", "11042", "11132"))

save(ThermRNAseqLow, file="Data/ThermRNAseqLow.rda")

#Thermal Tolerance High
ThermRNAseqHigh <- T_TDATA %>% filter(patRIL %in% c("12069", "12062", "11024", "11163", "11023", "11043", "11067"))

save(ThermRNAseqHigh, file="Data/ThermRNAseqHigh.rda")

```

##Transform data set (only if necessary). There are several methods used to transform data, choose the best fit.

```{r}

#log transformation, log10 transformation, arcsign transformation 


#sqaure root transformation
T_TDATA$Tolsqrtvariable<-sqrt(T_TDATA$Incapacitation_Mean)

#after transformation always visualize data
ggplot(T_TDATA, aes(x = T_TDATA$Tolsqrtvariable)) +
  geom_histogram()


#save new dataset
save(T_TDATA, file="../Data/T_TDATA.rda")


#power transformation

L_MDATA$LearnPowerTrans <- L_MDATA$Learning_Mean^3

ggplot(L_MDATA, aes(x = LearnPowerTrans)) +
  geom_histogram()

save(L_MDATA, file="../Data/L_MDATA.rda")

```

